Predicting Diabetes Based on Various Factors Using Machine Learning Models

### 2023-12-05 manuscript/manuscript.tex
Pruned Tree
Decision Tree continued

\section{Methods}
\label{sec:meth}
    Various methods will be used to complete the statistical analysis necessary for this project. I will begin by exploring basic 
    descriptive statistics to gain an understanding of my data set. I will then begin to create my model for my decision tree by
    splitting the data into a test and training set. The training set will encompass 80 percent of the data whereas the test set will
    only have 20 percent. After creating these two sets we will now be ready to build the model, this will be done by using the rpart
    function in R, where method will be 'class' as this is a binary model. We will then see in Figure~\ref{fig:structure} the structure
    of the decision tree created. Looking at the leftmost path in the decision tree we can see that at the top the outcome was 0 meaning 
    that the patient did not have diabetes 35 percent of the time. Moving down to the next node, we can see that 68 percent of patients with
    glucose level less than 128, did not have diabetes 19 percent of the time. To finish this path, 61 percent of the time if the patient was
    older than 29 years old, then 8 percent of the time the patient had diabetes. The rest of the structure is similar but with different 
    values and predictors.

@@

>>>>

\section{Methods}
\label{sec:meth}
    Various methods will be used to complete the statistical analysis necessary for this paper. I will begin by exploring basic 
    descriptive statistics to gain an understanding of my data set. We will move on to model testing, where we will perform both 
    decision tree analysis and random forest in order to determine what predictors prove to be necessary in predicting if a patient
    will have the outcome 1 of having diabetes. Decision trees were choosen to represent the data as they are generally pretty easy
    to understand.

    To begin, I will provide an explanation of what a decision tree is and why it is useful in this case. A decision tree is considered
    to be example of supervised learning used for regression and classification. Through our analysis we will present trees, these trees
    begin at the top with a decision root node, which is the point where you make a decision. Following down there are subsequent sub trees
    that have other decision nodes. Finally at the bottom of each tree you reach a leaf node, where it is the output of the decision and 
    there are no further branches off of this.

    I will then begin to create my model for my decision tree by
    splitting the data into a test and training set. The training set will encompass 80 percent of the data whereas the test set will
    only have 20 percent. After creating these two sets we will now be ready to build the model, this will be done by using the rpart
    function in R, where method will be 'class' as this is a binary model. We will then see in Figure~\ref{fig:structure} the structure
    of the decision tree created. Looking at the leftmost path in the decision tree we can see that at the top the outcome was 0 meaning 
    that the patient did not have diabetes 35 percent of the time. Moving down to the next node, we can see that 68 percent of patients with
    glucose level less than 128, did not have diabetes 19 percent of the time. To finish this path, 61 percent of the time if the patient was
    older than 29 years old, then 8 percent of the time the patient had diabetes. The rest of the structure is similar but with different 
    values and predictors.

@@

\begin{figure}[tbp]
  \centering
  \caption{Pruned Decision Tree Structure}
  \includegraphics[width=\textwidth]{Pruned_tree.pdf}
  \label{fig:prstructure}
  \end{figure}
